### 2023-02-02 Code/Supplement 1A_figures.Rmd

---
title: "Supplement 1A"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(tidyverse)
library(tidytext)
library(ggrepel)

library(gridExtra)

library(scoringutils)
library(viridis)

library(ggeffects)
library(emmeans)
library(splines)

setwd("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts")

### Need functions
source("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/Inc_Case_Eval_Manuscript/coverage.R") #coverage
### Estimate Relative WIS 
pairWIS_fun <-function(x, df) {

  df <-df %>%
    pairwise_comparison(., summarise_by = c("model"),
                        baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill) %>%
    distinct()
}
pairWIS_co_fun <-function(x, df) {

  df <-df %>%
    pairwise_comparison(., summarise_by = c("model", "pop_size_quant2"),
                        baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, pop_size_quant2, scaled_rel_skill) %>%
    distinct()
}

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size=10, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=11, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

## Function for setting up legends in empty facet spaces
library(gtable)
library(cowplot)

shift_legend <- function(p){

  # check if p is a valid object
  if(!"gtable" %in% class(p)){
    if("ggplot" %in% class(p)){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }

  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }

  # establish extent of unfilled facet panels (including any axis cells in between)
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
  names(empty.facet.panels) <- c("t", "l", "b", "r")

  # extract legend & copy over to location of unfilled facet panels
  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
  if(length(guide.grob) == 0){
    message("There is no legend present. Returning original plot.")
    return(p)
  }
  gp <- gtable_add_grob(x = gp,
                        grobs = gp[["grobs"]][[guide.grob]],
                        t = empty.facet.panels[["t"]],
                        l = empty.facet.panels[["l"]],
                        b = empty.facet.panels[["b"]],
                        r = empty.facet.panels[["r"]],
                        name = "new-guide-box")

  # squash the original guide box's row / column (whichever applicable)
  # & empty its cell
  guide.grob <- gp[["layout"]][guide.grob, ]
  if(guide.grob[["l"]] == guide.grob[["r"]]){
    gp <- gtable_squash_cols(gp, cols = guide.grob[["l"]])
  }
  if(guide.grob[["t"]] == guide.grob[["b"]]){
    gp <- gtable_squash_rows(gp, rows = guide.grob[["t"]])
  }
  gp <- gtable_remove_grobs(gp, "guide-box")

  return(gp)
}

## Change labels for counties
 rename_counts <-function(df) {   
   df <-df %>%
     mutate(pop_size_quant2=ifelse(pop_size_quant2==1, "County Forecasts: Size Quantile 1",
                                 ifelse(pop_size_quant2==2, "County Forecasts: Size Quantile 2",
                                        ifelse(pop_size_quant2==3, "County Forecasts: Size Quantile 3", 
                                               ifelse(pop_size_quant2==4, "County Forecasts: Size Quantile 4", "County forecasts: Size Quantile 5")))))
}

```

## Supplemental Figures

### Supplement 1

```{r Fig S1A, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=11}
start_date <-as.Date('2020-07-28')
stop_date <-as.Date('2021-12-21')

sec_models <-c("CU-nochange",  "CU-scenario_high", "CU-scenario_low", "CU-scenario_mid", "COVIDhub_CDC-ensemble") ## Non-primary models should be exlcuded

## Must have 4 weeks of forecasts for all locations
include_wks <-function(x, df){

  df <-df %>%
    group_by(model, sub_date, location) %>%
    mutate(no_horizons=length(unique(target))) %>% 
    ungroup() %>%
    filter(no_horizons>=4) #There should be at least 4 horizons
}

## Must have quantities for all forecasts
include_quant <-function(x, df){

  df <-df %>%
    group_by(model, location, forecast_date, target) %>%
    mutate(quants=length(quantile[!is.na(quantile)])) %>%
    ungroup() %>%
    filter(quants==7) #There should be 7 quantiles

}

## Must have forecasts for at least 50 locations for states/territories/district and national
include_locs <-function(x, df){

  df <-df %>%
    mutate(location=as.factor(location)) %>%
    group_by(model, forecast_date) %>%
    mutate(locs=length(unique(location))) %>% #[quantile==0.5 & target=="1 wk ahead inc case"]))) %>%
    ungroup() %>%
    filter(locs>=50)

}

## Must have 75% of all counties per quantile per submission date forecasted
include_locs_co <-function(x, df){

  df <-df %>%
    group_by(pop_size_quant2) %>%
    mutate(locs_total=length(unique(location))) %>% 
    ungroup() %>%
    group_by(model, pop_size_quant2, sub_date) %>%
    mutate(locs=length(unique(location))) %>% #[quantile==0.5 & target=="1 wk ahead inc case"]))) %>%
    ungroup() %>%
    mutate(percent_co=locs/locs_total) %>%
    filter(percent_co>=0.75)
}

## Must have submitted at least 50% of weeks
include_subs <-function(x, df){

  df <-df %>%
    filter(sub_date >= start_date) %>%
    mutate(total=length(unique(sub_date))) %>% 
    group_by(model) %>%
    mutate(subs=length(unique(sub_date))) %>%
    ungroup() %>%
    mutate(percent_sub=subs/total) %>%
    filter(percent_sub>=0.50)

}

############################################################################
############################################################################

load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/updated_cases_nat_state_forecasts_2022-04-04.Rdata")

### National + state submissions
nat_st <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  filter(type=="point")

### National + state submissions that met inclusion criteria
teams_nat_st <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  include_quant(df=.) %>% #all quantiles present
  include_wks(df=.) %>% #all targets included 
  include_locs(df=.) %>% #at least 50 locations
  include_subs(df=.) %>% #submitted at least 50% of the time
  dplyr::select(model) %>%
  mutate(model_rename=paste0("*", model)) %>%
  distinct()

### Submission plots

nat_st %>%
  dplyr::select(-target, -target_end_date, -value) %>%
  distinct() %>%
  group_by(sub_date, model) %>%
  summarize(locs=n()) %>%
  ungroup() %>%
  mutate(log_loc=log(locs)) %>%
  ungroup() %>%
  mutate(total_sub=length(unique(sub_date))) %>%
  group_by(model) %>%
  mutate(wks_sub=length(sub_date),
         per_sub=wks_sub/total_sub) %>%
  ungroup() %>%
  left_join(., teams_nat_st, by="model") %>%
  mutate(model_rename=ifelse(is.na(model_rename), model, model_rename),
         model_rename=fct_reorder(model_rename, wks_sub)) %>%
  ggplot(., aes(sub_date, model_rename)) +
  geom_tile(aes(fill=log_loc)) +
  scale_fill_gradient(low = "#c7e9c0", high = "#41ab5d") + 
  geom_text(aes(label=locs), size=1.50) +
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B") +
  theme_bw() +
  labs(title = "A. Primary Model* Submissions for National/State/Territory-Level Case Forecasts Over Time",
       subtitle= "The number of locations forecasted per submission is labled",
       caption = "*Model met inclusion criteria") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        legend.position = "none")

```