### 2023-02-02 Code/Supplement 1B_figures.Rmd

---
title: "Supplement 1.B"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(tidyverse)
library(tidytext)
library(ggrepel)

library(gridExtra)

library(scoringutils)

setwd("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts")

### Need functions
source("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/Inc_Case_Eval_Manuscript/coverage.R") #coverage

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size=10, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=11, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

## Change labels for counties
 rename_counts <-function(df) {   
   df <-df %>%
     mutate(pop_size_quant2=ifelse(pop_size_quant2==1, "County Forecasts: Size Quantile 1",
                                 ifelse(pop_size_quant2==2, "County Forecasts: Size Quantile 2",
                                        ifelse(pop_size_quant2==3, "County Forecasts: Size Quantile 3", 
                                               ifelse(pop_size_quant2==4, "County Forecasts: Size Quantile 4", "County forecasts: Size Quantile 5")))))
}

```

## Supplemental Figures

### Supplement 1

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=11}
start_date <-as.Date('2020-07-28')
stop_date <-as.Date('2021-12-21')

sec_models <-c("CU-nochange",  "CU-scenario_high", "CU-scenario_low", "CU-scenario_mid", "COVIDhub_CDC-ensemble") ## Non-primary models should be exlcuded

## Must have 4 weeks of forecasts for all locations
include_wks <-function(x, df){

  df <-df %>%
    group_by(model, sub_date, location) %>%
    mutate(no_horizons=length(unique(target))) %>% 
    ungroup() %>%
    filter(no_horizons>=4) #There should be at least 4 horizons
}

## Must have quantities for all forecasts
include_quant <-function(x, df){

  df <-df %>%
    group_by(model, location, forecast_date, target) %>%
    mutate(quants=length(quantile[!is.na(quantile)])) %>%
    ungroup() %>%
    filter(quants==7) #There should be 7 quantiles

}

## Must have 75% of all counties per quantile per submission date forecasted
include_locs_co <-function(x, df){

  df <-df %>%
    group_by(pop_size_quant2) %>%
    mutate(locs_total=length(unique(location))) %>% 
    ungroup() %>%
    group_by(model, pop_size_quant2, sub_date) %>%
    mutate(locs=length(unique(location))) %>% 
    ungroup() %>%
    mutate(percent_co=locs/locs_total) %>%
    filter(percent_co>=0.75)
}

## Must have submitted at least 50% of weeks
include_subs <-function(x, df){

  df <-df %>%
    filter(sub_date >= start_date) %>%
    mutate(total=length(unique(sub_date))) %>% 
    group_by(model) %>%
    mutate(subs=length(unique(sub_date))) %>%
    ungroup() %>%
    mutate(percent_sub=subs/total) %>%
    filter(percent_sub>=0.50)

}

############################################################################
############################################################################

```

```{r Fig S1B, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=11}

####
load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/cases_county_forecasts_2022-04-04.Rdata")
size <-read.csv("C:/Users/oko8/Downloads/SVI2018_US_COUNTY.csv") %>%
  dplyr::select(FIPS, E_TOTPOP) %>%
  rename(location=FIPS) %>%
  mutate(location=as.character(location),
         location=ifelse(nchar(location)==4, paste0("0", location), location),
         pop_size_quant2=cut(E_TOTPOP, breaks = quantile(E_TOTPOP, probs = seq(0, 1, 0.20)),
                             include.lowest = TRUE, labels = 1:5)
  ) %>% 
  dplyr::select(-E_TOTPOP) 
all_dat <-all_dat %>% left_join(., size, by="location") 
rm(size)
counts <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date >= start_date & sub_date <= stop_date) %>%
  filter(type=="point")

# Inclusion criteria for counties based on pop size
all_dat <-all_dat %>%
  filter(!(model %in% sec_models),
         sub_date <= stop_date) 
all_dat <-all_dat %>% include_quant(df=.) 
all_dat <-all_dat %>% include_wks(df=.) 
all_dat <-all_dat %>% include_subs(df=.) 
all_dat <-all_dat %>% include_locs_co(df=.)

all_dat <-all_dat %>%
  dplyr::select(-quants, -no_horizons, -locs, -locs_total, -percent_co, -total, -subs, -percent_sub) %>%
  dplyr::select(model) %>%
  mutate(model_rename=paste0("*", model)) %>%
  distinct()

counts %>%
  dplyr::select(-target, -target_end_date, -value) %>%
  distinct() %>%
  group_by(sub_date, model, pop_size_quant2) %>%
  summarize(locs=n()) %>%
  rename_counts() %>%
  ungroup() %>%
  mutate(log_loc=log(locs)) %>%
  mutate(total_sub=length(unique(sub_date))) %>%
  ungroup() %>%
  group_by(model, pop_size_quant2) %>%
  mutate(wks_sub=length(sub_date),
         per_sub=wks_sub/total_sub) %>%
  ungroup() %>%
  left_join(., all_dat, by="model") %>%
  mutate(model_rename=ifelse(is.na(model_rename), model, model_rename),
         model_rename=fct_reorder(model_rename, wks_sub)) %>%
  ggplot(., aes(sub_date, model_rename)) +
  geom_tile(aes(fill=per_sub)) +
  scale_fill_gradient(low = "#c7e9c0", high = "#4a1486",
                      name="Percent of \nCounties \nForecasted") + 
  scale_x_date(date_breaks = "1 month", date_minor_breaks = "1 week", date_labels = "%B") +
  facet_wrap(~ pop_size_quant2) +
  theme_bw() +
  labs(title = "B. Primary Model* Submissions for County-Level Case Forecasts Over Time",
       caption = "*Model met inclusion criteria") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        strip.background = strip.background_standard, strip.text = strip.text_standard,
        legend.position = "bottom")

```