### 2023-02-02 Code/Supplement 3.2 and 3.3_figures.rmd

---
title: "Supplement 3.2 and 3.3"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(tidyverse)
library(tidytext)
library(ggrepel)

library(gridExtra)
library(cowplot)
library(scoringutils)
library(viridis)

library(ggeffects)
library(emmeans)
library(splines)

setwd("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts")

### Need functions
source("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/Inc_Case_Eval_Manuscript/coverage.R") #coverage

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size=10, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=11, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

## Change labels for counties
 rename_counts <-function(df) {   
   df <-df %>%
     mutate(pop_size_quant2=ifelse(pop_size_quant2==1, "County Forecasts: Size Quantile 1",
                                 ifelse(pop_size_quant2==2, "County Forecasts: Size Quantile 2",
                                        ifelse(pop_size_quant2==3, "County Forecasts: Size Quantile 3", 
                                               ifelse(pop_size_quant2==4, "County Forecasts: Size Quantile 4", "County forecasts: Size Quantile 5")))))
}

 stop_date <-as.Date("2021-12-21")
```

## Supplemental Figures

### Supplement 3: Non large country figures

#### Absolute WIS

```{r Fig S3.3, echo=FALSE, warning=FALSE, message=FALSE, fig.height=11, fig.width=12}

# Function for WIS over time
WIS.ot.fig_dat <-function(df) {

  sub_date <-c(unique(df$sub_date))
  model <-c(as.character(unique(df$model)))
  model_dates <-crossing(sub_date, model)

  cast_dates <-df %>%
    dplyr::select(sub_date, date, pop_size_quant2) %>%
    distinct() %>%
    left_join(model_dates, by="sub_date")

  dat <-df %>%
    full_join(cast_dates, ., by=c("model", "sub_date", "date", "pop_size_quant2")) %>%
    mutate(bold=ifelse(model=="COVIDhub-baseline" | model=="COVIDhub-4_week_ensemble", "T", "F")) %>%
    group_by(model, date, bold, pop_size_quant2) %>%
    summarize(abWIS=mean(interval_score, na.rm=TRUE)) %>%
    ungroup() %>%
    mutate(grouping=ifelse(model=="COVIDhub-baseline", "COVIDhub-baseline", 
                           ifelse(model=="COVIDhub-4_week_ensemble", "COVIDhub-4_week_ensemble", "Other teams"))) #%>%
  return(dat)

}
cov.ot.fig_dat <-function(df) {

  sub_date <-c(unique(df$sub_date))
  model <-c(as.character(unique(df$model)))
  model_dates <-crossing(sub_date, model)

  cov <-df %>%
    group_by(model, target_end_date, pop_size_quant2) %>%
    coverage(df=.) %>%
    summarize(cov.95=mean(coverage.95))

  dat <-df %>%
    dplyr::select(sub_date, target_end_date, pop_size_quant2) %>%
    distinct() %>%
    left_join(., model_dates, by=c("sub_date")) %>%
    dplyr::select(-sub_date) %>%
    left_join(., cov, by=c("model", "target_end_date", "pop_size_quant2")) %>%
    distinct() %>%
    ungroup() %>%
    mutate(grouping=ifelse(model=="COVIDhub-baseline", "COVIDhub-baseline", 
                           ifelse(model=="COVIDhub-4_week_ensemble", "COVIDhub-4_week_ensemble", "Other teams"))) %>%
    mutate(bold=ifelse(model=="COVIDhub-baseline" | model=="COVIDhub-4_week_ensemble", "T", "F"))

  return(dat)

}
ot.plot <-function(df, x, y) {

  ggplot() +
    geom_line(data=filter(df, bold=="F"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    geom_point(data=filter(df, bold=="F"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    geom_line(data=filter(df, bold=="T"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    geom_point(data=filter(df, bold=="T"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    scale_size_manual(values = c("T"=1.25, "F"=0.85)) +
    scale_x_date(date_labels="%b-%y", breaks = df$x, date_breaks ="3 months", date_minor_breaks ="1 month") +
    geom_vline(xintercept = as.numeric(stop_date), linetype=4, size=1.05) +
    scale_colour_manual(name="",
                        values = c("COVIDhub-baseline"="#762a83",
                                   "COVIDhub-4_week_ensemble"="#1b7837",
                                   "Other teams"="grey80"),
                        labels=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble"),
                        breaks=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble")) +
    guides(size = "none") + 
    theme_bw() +
    theme(legend.position = "none", 
          axis.title.x = element_blank())
}

###########################################################################
############################################################################

### WIS Over time
load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/WIS_all horizons_non large counties_all subs.rdata")

## National
WIS_ot <-WIS.ot.fig_dat(WIS_all) %>%
  rename_counts() #%>%
  #mutate(abWIS=ifelse(abWIS=="NaN", NA, abWIS))
WIS <-ot.plot(WIS_ot, date, abWIS)  +
  facet_wrap(~ pop_size_quant2, ncol = 1) +
  scale_y_continuous(trans='log10') +
  theme(axis.title.x=element_blank(),
        legend.position = c(0.25, 0.94),legend.title = element_blank(),
        legend.justification = c(0, 0.5), 
        legend.margin=margin(c(0.025,0.025,0.025,0.025)),
        legend.key = element_rect(colour = NA, fill = NA), 
        strip.background = strip.background_standard, strip.text = strip.text_standard) +
  labs(title="A. Average Weighted Interval Score for County Forecasts",
       x = "Forecast target date",
       y = "Average Weighted Interval Score")

############################################################################
############################################################################

### Coverage Over time
load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/cases_non large counties_forecasts_for analysis_2022-08-18.Rdata")

cov_ot <-cov.ot.fig_dat(dat_other_count) %>%
  rename_counts() 
cov <-ot.plot(cov_ot, target_end_date, cov.95)  +
  geom_hline(yintercept = 0.95,  linetype=6, size=1.05) +
  facet_wrap(~ pop_size_quant2, ncol = 1) +
  #theme(axis.title.x=element_blank()) +  
  labs(title="B. 95% Coverage for County Forecasts",
       x = "Forecast target date",
       y = "95% Coverage") + 
  theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5),
        strip.background = strip.background_standard, strip.text = strip.text_standard)

############################################################################
############################################################################

grid.arrange(WIS, cov,
             ncol=2)

```

####  Coverage.

```{r Fig S3.2, echo=FALSE, warning=F, message=F, fig.height=8.5, fig.width=9}

# Read in and manipulate data + plot structure
cov.fig_dat <-function(df) {
  coverage <-df %>%
    coverage(df=.) %>%
    group_by(model, pop_size_quant2) %>%
    summarize(coverage.95=mean(coverage.95),
              coverage.80=mean(coverage.80),
              coverage.50=mean(coverage.50)) %>%
    gather("expected_cov", "observed_cov", coverage.95:coverage.50) %>%
    mutate(expected_cov=gsub("coverage.", "", expected_cov),
           expected_cov=as.numeric(expected_cov),
           expected_cov=expected_cov/100,
           bold=ifelse(model=="COVIDhub-baseline" | model=="COVIDhub-4_week_ensemble", "T", "F"),
          grouping=ifelse(model=="COVIDhub-baseline", "COVIDhub-baseline", 
                           ifelse(model=="COVIDhub-4_week_ensemble", "COVIDhub-4_week_ensemble", "Other teams")))
  return(coverage)
}
data_labels <-function(df) {

  ensemble_cov <-df%>%
    filter(model=="COVIDhub-4_week_ensemble")

  data_labels <-df %>%  
    filter(model!="COVIDhub-4_week_ensemble") %>%
    mutate(test_95=ifelse(expected_cov==0.95 & observed_cov > ensemble_cov$observed_cov[ensemble_cov$expected_cov==0.95], 1, NA),
           test_80=ifelse(expected_cov==0.80 & observed_cov > ensemble_cov$observed_cov[ensemble_cov$expected_cov==0.80], 1, NA),
           test_50=ifelse(expected_cov==0.50 & observed_cov > ensemble_cov$observed_cov[ensemble_cov$expected_cov==0.50], 1, NA)) %>%
    filter(!is.na(test_95) | !is.na(test_80) | !is.na(test_50)) %>%
    group_by(model) %>%
    mutate(model_count=n()) %>%
    filter(model_count==3 & expected_cov==0.95) %>%
    mutate(model=ifelse(model=="JHU_UNC_GAS-StatMechPool", "JHU_UNC_GAS\n-StatMechPool", model)) %>%
    mutate(model=ifelse(model=="MOBS-GLEAM_COVID", "MOBS-GLEAM_\nCOVID", model)) %>%
    mutate(model=ifelse(model=="IEM_MED-CovidProject", "IEM_MED\n-CovidProject", model)) %>%
    mutate(model=ifelse(model=="COVIDhub-trained_ensemble", "COVIDhub\n-trained\n_ensemble", model)) %>%
  return(data_labels)
  }
cov.plot <-function(df, df2) {
  ggplot(df) +
    geom_line(data=filter(df, bold=="F"), 
              mapping=aes(expected_cov, observed_cov, group=model, color=grouping, size=bold, alpha=0.75)) +
    geom_point(data=filter(df, bold=="F"), 
               mapping=aes(expected_cov, observed_cov, group=model, color=grouping), size=1.5) +
    geom_line(data=filter(df, bold=="T"), 
              mapping=aes(expected_cov, observed_cov, group=model, color=grouping, size=bold, alpha=0.75)) +
    geom_point(data=filter(df, bold=="T"), 
               mapping=aes(expected_cov, observed_cov, group=model, color=grouping), size=3) +
    geom_line(aes(expected_cov, expected_cov, color=grouping, group=model), linetype="dashed", size=1.5, color="black") +
    scale_size_manual(values = c("T"=1.5, "F"=0.85)) +
    scale_colour_manual(values = c("COVIDhub-baseline"="#762a83",
                                   "COVIDhub-4_week_ensemble"="#1b7837",
                                   "Other teams"="grey70"),
                        labels=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week\n_ensemble"),
                        breaks=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble")) +
    guides(size = "none", alpha="none") +
    labs(y="Observed Coverage") +
    theme_bw()  + 
    theme(legend.position = "none", 
          axis.title.x = element_blank()) +
    geom_text_repel(data=df2,
                    mapping=aes(expected_cov, observed_cov, label=model), size=3,
                    force=1, point.padding=unit(0.5,'lines'), nudge_x =0.1,
                    segment.size= 0.05,  segment.color="grey50",
                    direction="y", #hjust = "right",
                    vjust=0.5) +
    scale_x_continuous(expand = expansion(mult = c(0.05, 0.25)))
}

# Figures

# All Counties
coverage_count1 <-cov.fig_dat(filter(dat_other_count, pop_size_quant2==1, sub_date <= stop_date))
labels_count1 <-data_labels(coverage_count1)

p1 <-cov.plot(coverage_count1, labels_count1) +
  labs(title="County Forecasts: Size Quantile 1") +
  theme(legend.position = c(0.02, 0.88), legend.title = element_blank(),
        legend.justification = c(0, 0.5), 
        legend.margin=margin(c(0.5,1,1,1)),
        legend.key = element_rect(colour = NA, fill = NA))

coverage_count2 <-cov.fig_dat(filter(dat_other_count, pop_size_quant2==2, sub_date <= stop_date))
labels_count2 <-data_labels(coverage_count2) 
p2 <-cov.plot(coverage_count2, labels_count2) +
  labs(title="County Forecasts: Size Quantile 2", 
       y="")

coverage_count3 <-cov.fig_dat(filter(dat_other_count, pop_size_quant2==3, sub_date <= stop_date))
labels_count3 <-data_labels(coverage_count3) 
p3 <-cov.plot(coverage_count3, labels_count3) +
  labs(title="County Forecasts: Size Quantile 3",
       x="Expected Coverage")

coverage_count4 <-cov.fig_dat(filter(dat_other_count, pop_size_quant2==4, sub_date <= stop_date))
labels_count4 <-data_labels(coverage_count4) 
p4 <-cov.plot(coverage_count4, labels_count4) +
  labs(title="County Forecasts: Size Quantile 4",
       y="",
       x="Expected Coverage")

grid.arrange(p1, p2, p3, p4,
              ncol=2)

```