### 2023-02-02 Code/Supplement 6_figures.Rmd

---
title: "Supplement 6"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(tidyverse)
library(tidytext)
library(ggrepel)

library(gridExtra)

library(scoringutils)
library(viridis)

library(ggeffects)
library(emmeans)
library(splines)

library(grid)

setwd("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts")

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size=10, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=11, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

## Function for setting up legends in empty facet spaces
library(gtable)
library(cowplot)

shift_legend <- function(p){

  # check if p is a valid object
  if(!"gtable" %in% class(p)){
    if("ggplot" %in% class(p)){
      gp <- ggplotGrob(p) # convert to grob
    } else {
      message("This is neither a ggplot object nor a grob generated from ggplotGrob. Returning original plot.")
      return(p)
    }
  } else {
    gp <- p
  }

  # check for unfilled facet panels
  facet.panels <- grep("^panel", gp[["layout"]][["name"]])
  empty.facet.panels <- sapply(facet.panels, function(i) "zeroGrob" %in% class(gp[["grobs"]][[i]]))
  empty.facet.panels <- facet.panels[empty.facet.panels]
  if(length(empty.facet.panels) == 0){
    message("There are no unfilled facet panels to shift legend into. Returning original plot.")
    return(p)
  }

  # establish extent of unfilled facet panels (including any axis cells in between)
  empty.facet.panels <- gp[["layout"]][empty.facet.panels, ]
  empty.facet.panels <- list(min(empty.facet.panels[["t"]]), min(empty.facet.panels[["l"]]),
                             max(empty.facet.panels[["b"]]), max(empty.facet.panels[["r"]]))
  names(empty.facet.panels) <- c("t", "l", "b", "r")

  # extract legend & copy over to location of unfilled facet panels
  guide.grob <- which(gp[["layout"]][["name"]] == "guide-box")
  if(length(guide.grob) == 0){
    message("There is no legend present. Returning original plot.")
    return(p)
  }
  gp <- gtable_add_grob(x = gp,
                        grobs = gp[["grobs"]][[guide.grob]],
                        t = empty.facet.panels[["t"]],
                        l = empty.facet.panels[["l"]],
                        b = empty.facet.panels[["b"]],
                        r = empty.facet.panels[["r"]],
                        name = "new-guide-box")

  # squash the original guide box's row / column (whichever applicable)
  # & empty its cell
  guide.grob <- gp[["layout"]][guide.grob, ]
  if(guide.grob[["l"]] == guide.grob[["r"]]){
    gp <- gtable_squash_cols(gp, cols = guide.grob[["l"]])
  }
  if(guide.grob[["t"]] == guide.grob[["b"]]){
    gp <- gtable_squash_rows(gp, rows = guide.grob[["t"]])
  }
  gp <- gtable_remove_grobs(gp, "guide-box")

  return(gp)
}

```

## Supplemental Figures

### Supplement 6

```{r echo=FALSE, warning=F, message=F}

load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Data/WIS_gee.Rdata") # data
load("C:/Users/oko8/OneDrive - CDC/COVID/Forecasts/Code/model_caseslogstd_noint_gee.rda")

WIS_gee <-WIS_gee %>%
  group_by(model) %>%
  mutate(WIS_log_std=(log(interval_score)-mean(log(interval_score)))/sd(log(interval_score)),
         cases_log_std=(log(true_value)-mean(log(true_value)))/sd(log(true_value))
  ) %>%
  ungroup()

team <-c(unique(as.character(WIS_gee$model))) 
dat_team <-list()
means_sds <-list()
marg_means <-list()

# Model Predictions 
for(i in 1:length(team)){

  dat_team[[i]] <-as.data.frame(WIS_gee %>%
                                  filter(model==team[i]) %>%
                                  mutate(model=droplevels(model)) %>%
                                  ungroup())

  means_sds[[i]] <-dat_team[[i]] %>%
    mutate(mean_log_WIS=mean(log(interval_score)),
           sd_log_WIS=sd(log(interval_score)),
           mean_log_cases=mean(log(true_value)),
           sd_log_cases=sd(log(true_value))
    ) %>%
    dplyr::select(model, contains("mean_"), contains("sd_")) %>%
    distinct()

  marg_means[[i]] <-ggemmeans(model[[i]], terms=c("phase_all", "cases_log_std[-1.96:1.96]")) %>%
    rename(phase_all=x,
           cases=group) %>% 
    mutate(model=team[i],
           predicted_back=(predicted*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_lo=(conf.low*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_up=(conf.high*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           cases=as.numeric(as.character(cases)),
           cases_log_back=(cases*means_sds[[i]]$sd_log_cases) + means_sds[[i]]$mean_log_cases) %>%
    mutate(predicted_back_rescale=exp(predicted_back),
           predicted_back_lo_rescale=exp(predicted_back_lo),
           predicted_back_up_rescale=exp(predicted_back_up),
           cases_back_rescale=exp(cases_log_back))
}
marg_means <-do.call(rbind.data.frame, marg_means)

marg_means <-marg_means %>%

  mutate(model=as.factor(model)) %>%
  mutate(phase_all=as.factor(ifelse(phase_all=="increasing", "Increasing",
                                    ifelse(phase_all=="decreasing", "Decreasing",
                                           ifelse(phase_all=="peak", "Peak", "Nadir")))),
         phase_all=factor(phase_all, 
                          levels = c("Increasing", "Peak", "Decreasing",  "Nadir"))
  )

# Figure
FigS6 <-marg_means %>%
  ggplot(., aes(cases_back_rescale, predicted_back_rescale)) +
  geom_line(aes(color=phase_all)) +
  geom_ribbon(aes(ymin=predicted_back_lo_rescale, ymax=predicted_back_up_rescale, fill=phase_all), alpha=0.35) +
  scale_fill_manual(breaks = c("Increasing", "Decreasing", "Unclassified", "Peak", "Nadir"), 
                      values=c("#fdae61", "#abdda4", "#bababa", "#d7191c", "#2b83ba")) +
  scale_color_manual(breaks = c("Increasing", "Decreasing", "Unclassified", "Peak", "Nadir"), 
                      values=c("#fdae61", "#abdda4", "#bababa", "#d7191c", "#2b83ba")) +
  labs( y="Marginal Mean \nWeighted Interval Score", x="Reported cases") +
  facet_wrap(~model) +
  theme_bw() +
  theme(axis.title.y=axis.title.y_standard,
        legend.title = element_blank(), legend.text=legend.text_standard,
        strip.text=strip.text_standard, strip.background=strip.background_standard,
        axis.text.x = element_text(angle = 45, hjust=1)) +
  guides(fill=guide_legend(nrow=2, byrow=TRUE))

```

```{r Fig s6, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=11.5}

grid.draw(shift_legend(FigS6))

```