### 2023-04-18 Code/Manuscript_figures.Rmd
cleaned up code

---
title: "Inc COVID-19 Case Forecast Eval: Results"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(here)
library(tidyverse)
library(tidytext)
library(ggrepel)

library(gridExtra)
library(cowplot)
library(ggpubr)
library(scales)

library(scoringutils)
library(viridis)

library(ggeffects)
library(emmeans)
library(splines)

cut_date <-as.Date("2021-12-21")

### Need functions
source(here("Code/coverage.R")) #coverage

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size = 12, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=10, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

```

## Main Text Figures

### Figure 1. Select ensemble forecasts

```{r Fig1, echo=FALSE, warning=F, message=F, fig.height=6.5, fig.width=6.5}
### Reported cases over time
source(here("Code/Reported cases.R")) #reported cases
obs_data <-rd_obs() %>%
  filter(nchar(location) < 3) %>%
  filter(date <= (cut_date+4)+(7*3)) %>%
  mutate(true_value=ifelse(true_value <=0, NA, true_value),
         true_value.pop=ifelse(is.na(true_value), NA, true_value.pop))

Fig1.A <-ggplot() +
  geom_line(data=filter(obs_data, grouping=="State/Territory/DC"), 
            mapping=(aes(date, true_value.pop, group=location_name, color=grouping)), size=0.85) +
  scale_x_date(date_labels="%b-%y", breaks = obs_data$date, date_breaks ="3 months", date_minor_breaks ="1 month") +
  geom_line(data=filter(obs_data, grouping=="National"), 
            mapping=(aes(date, true_value.pop, group=location_name, color=grouping)), size=1.25) +
  geom_vline(xintercept = as.numeric(cut_date),  linetype=4, size=1.05) +
  scale_colour_manual(name="",
                      values = c("National"="black",
                                 "State/Territory/DC"="grey75")) +
  labs(title = "A. Reported Cases Per 100K Population",
       y = "Incident COVID-19 Cases per 100K population") +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        legend.position = c(0.02, 0.82),legend.title = element_blank(),
        legend.justification = c(0, 0.5), 
        legend.margin=margin(c(0.5,1,1,1)),
        legend.key = element_rect(colour = NA, fill = NA)) + theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5))

load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))

cast_dates <-dat_US_state %>% 
  filter(model=="COVIDhub-4_week_ensemble") %>% 
  ungroup() %>% 
  dplyr::select(sub_date) %>% 
  unique() %>% 
  arrange() %>%
  slice(which(row_number() %% 6 == 1))

Fig1.B <-dat_US_state %>%
  filter(sub_date <= cut_date) %>%
  filter(location=="US", 
         model=="COVIDhub-4_week_ensemble") %>% 
  filter(sub_date %in% cast_dates$sub_date) %>%
  mutate(value = case_when(quantile==0.5 ~ round(value),
                             quantile<0.5 ~ floor(value),
                             quantile>0.5 ~ ceiling(value),
                             type=='point' ~ round(value)), 
         model=as.factor(model), 
         model=droplevels(model)) %>% 
  pivot_wider(names_from = c(type, quantile), values_from=value) %>%
  rename(point = point_NA) %>% 
  ungroup() %>%
  ggplot(.) +
  geom_line(data=filter(obs_data, grouping=="National"), mapping=aes(x=date, y=true_value), color="grey85") +
  geom_point(data=filter(obs_data, grouping=="National"), mapping=aes(x=date, y=true_value), color="grey75") +
  geom_line(aes(target_end_date, point, group=interaction(sub_date, model), color=model), size=1.05) + 
  geom_point(aes(target_end_date, point, color=model), size=1.25) +
  geom_ribbon(aes(x=target_end_date, ymin=quantile_0.025, ymax=quantile_0.975,
                  fill=model, group=interaction(sub_date, model)), alpha=0.25) +
  geom_vline(xintercept = as.numeric(cut_date),  linetype=4, size=1.05) +
  scale_colour_manual(name="",
                      values = c("COVIDhub-4_week_ensemble"="#1b7837")) +
  scale_fill_manual(name="",
                    values = c("COVIDhub-4_week_ensemble"="#1b7837")) +
   scale_x_date(date_labels="%b-%y", breaks = obs_data$date, date_breaks ="3 months", date_minor_breaks ="1 month") +
  labs(title="B. Select National forecasts",
       y="Incident COVID-19 Cases") +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5),
        legend.title = element_blank(), legend.position = c(0.20, 0.82)) + scale_y_continuous(label=comma,
                                                                                               breaks=pretty_breaks(2))
plot_grid(Fig1.A, Fig1.B,
          nrow=2)

```

Figure 2. 95% coverage and relWIS  and submissions per team

```{r echo=FALSE, warning=FALSE, message=FALSE}

## WIS estimates

mean_WIS <-function() {
  load(here("Data/WIS_all horizons_nat_state.rdata"))

  ## National
  WIS_nat <-WIS_all %>%
    ungroup() %>%
    filter(location_name=="National") %>%
    summarise_scores(by = c("model"),
                     metric="interval_score",
                     relative_skill = TRUE,
                     baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill) %>%
    mutate(model_reorder=model) %>%
    rename(`National Forecasts`=scaled_rel_skill) 
  WIS_nat <-WIS_nat[, c(1, 3, 2)]

  ## State/Territory/District 
  WIS_state <-WIS_all %>%
    filter(location_name!="National") %>% 
    summarise_scores(by = c("model"),
                     metric="interval_score",
                     relative_skill = TRUE,
                     baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill) %>%
    rename(`State/Territory/DC Forecasts`=scaled_rel_skill) %>%
    mutate(model_reorder=fct_reorder(model, `State/Territory/DC Forecasts`))

  WIS_levels <-levels(WIS_state$model_reorder)

  mean_WIS <-WIS_state %>%

    full_join(., WIS_nat) %>% 
    mutate(model_reorder=factor(model, levels = WIS_levels)) %>%
    arrange(`State/Territory/DC Forecasts`) %>% 
    gather(key, value, -model, -model_reorder) %>%
    mutate(key=factor(key, levels = c("National Forecasts", 
                                      "State/Territory/DC Forecasts"))) %>%
    mutate(value_display=format(round(value, digits=2), nsmall = 2),
           value_display=ifelse(value_display=="  NA", NA, value_display))

  }
mean_WIS <-mean_WIS()
mean_WIS$model_reorder <-fct_rev(mean_WIS$model_reorder)
mean_WIS$key <-fct_rev(mean_WIS$key)
levels_model <-levels(mean_WIS$model_reorder)
levels_key <-levels(mean_WIS$key)

WIS.heatmap <-mean_WIS %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 1.0, na.value = "grey75") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="Relative Weighted Interval Score",
       title="B. Relative Weighted Interval Score") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(), axis.text.y=element_blank(),
        axis.text=axis.text_standard,
        legend.position="bottom")

############################################################################
############################################################################

### Coverage estimates

cov.95 <-function() {

  load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))

  coverage_US <- dat_US_state %>%
    filter(sub_date <= cut_date) %>%
    filter(location=="US") %>%
    coverage(df=.) %>%
    group_by(model) %>%
    summarize(`National Forecasts`=mean(coverage.95))

  ## State/Territory/District 
  coverage_state <- dat_US_state %>%
    filter(sub_date <= cut_date) %>%
    filter(location!="US") %>%
    coverage(df=.) %>%
    group_by(model) %>%
    summarize(`State/Territory/DC Forecasts`=mean(coverage.95))

  cov.95 <-coverage_US %>%
    full_join(., coverage_state, by="model") %>%
    gather(key, value, -model) %>%
    mutate(cov_diff=abs(0.95-value)) %>%
    mutate(value_display=format(round(value, digits=2), nsmall = 2),
           value_display=ifelse(value_display=="  NA", NA, value_display))
}

cov.95 <-cov.95() %>%
  mutate(model_reorder=factor(model, levels=levels_model), 
         key=factor(key, levels=levels_key))

cov.heatmap <-cov.95 %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=value)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient2(low = "white", mid = "#f7fcb9", high = "#31a354", midpoint = 0.75, na.value = "grey75") +
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="95% Coverage",
       title="C. 95% Coverage") +
  theme(axis.title.x = element_blank(), #axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(), axis.text.y=element_blank(),
        axis.text=axis.text_standard,
        legend.position="bottom")

############################################################################
############################################################################

### Submissions over time

submissions <-function(df) {

  start_date <-as.Date("2020-07-28")
  stop_date <-as.Date("2022-02-22")

  df <-df %>%
    filter(sub_date >= start_date & sub_date <= cut_date) %>% #stop_date) %>%
    dplyr::select(forecast_date, model, location, target, sub_date) %>%
    distinct() %>%
    dplyr::select(-target) %>%
    distinct() %>%
    group_by(sub_date, model) %>%
    summarize(locs=n()) %>%
    ungroup() %>%
    mutate(total_sub=length(unique(sub_date))) %>%
    group_by(model) %>%
    mutate(wks_sub=length(sub_date),
           per_sub=(wks_sub/total_sub)) %>%
    ungroup() %>%
  dplyr::select(-locs, -sub_date) %>%
  distinct()

  return(df)

}

all_dat <-function(){

  load(here("Data/cases_nat_state_forecasts_2021-03-13.Rdata"))
  load(here("Data/cases_nat_state_forecasts_2022-03-21.Rdata"))

  all_dat <-all_dat %>%
    bind_rows(., all_dat_updated) %>%
    mutate(model=ifelse(sub_date < "2021-03-16" & model=="COVIDhub-ensemble", "COVIDhub-4_week_ensemble", model))
}
all_dat <-all_dat()
subs_nat <-submissions(df=filter(all_dat, location=="US")) %>% 
  filter(model %in% mean_WIS$model[mean_WIS$key=="National Forecasts" & !is.na(mean_WIS$value)]) %>%
  mutate(key="National Forecasts")
subs_state <-submissions(df=filter(all_dat, location!="US")) %>% 
  filter(model %in% mean_WIS$model[mean_WIS$key=="State/Territory/DC Forecasts"]) %>%
  mutate(key="State/Territory/DC Forecasts")

missings <-mean_WIS %>%
  filter(is.na(value)) %>%
  rename(per_sub=value)

all_subs <-subs_nat %>%
  bind_rows(., subs_state) %>% 
  mutate(model_reorder=factor(model, levels=levels_model), 
         key=factor(key, levels=levels_key),
         value_display=format(round(per_sub, digits=2), nsmall = 2),
         value_display=ifelse(value_display=="  NA", NA, value_display)) %>%
  bind_rows(., missings) %>%
  ggplot(., aes(key, model_reorder)) +
  geom_tile(aes(fill=per_sub)) +
  geom_text(aes(label=value_display), size=3.5) +
  scale_fill_gradient(low = "#fde0dd", high = "#c51b8a", na.value = "grey75") +
  scale_x_discrete(labels = function(x) str_wrap(x, width=15)) +
  theme_bw() +
  labs(fill="% Weeks \nSubmitted",
       title="A. Percent of Weeks Submitted") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        axis.text=axis.text_standard,
        legend.position="bottom")

############################################################################
############################################################################

```

``` {r Fig 2, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=12}

plot_grid(all_subs, WIS.heatmap, cov.heatmap,
          nrow=1, 
          rel_widths = c(1, 0.6, 0.6))

```

### Figure 3. Coverage.

```{r echo=FALSE, warning=F, message=F, fig.show='hide'}

# Read in and manipulate data + plot structure
cov.fig_dat <-function(df) {
  coverage <-df %>%
    coverage(df=.) %>%
    group_by(model) %>%
    summarize(coverage.95=mean(coverage.95),
              coverage.80=mean(coverage.80),
              coverage.50=mean(coverage.50)) %>%
    gather("expected_cov", "observed_cov", coverage.95:coverage.50) %>%
    mutate(expected_cov=gsub("coverage.", "", expected_cov),
           expected_cov=as.numeric(expected_cov),
           expected_cov=expected_cov/100,
           bold=ifelse(model=="COVIDhub-baseline" | model=="COVIDhub-4_week_ensemble", "T", "F"),
          grouping=ifelse(model=="COVIDhub-baseline", "COVIDhub-baseline", 
                           ifelse(model=="COVIDhub-4_week_ensemble", "COVIDhub-4_week_ensemble", "Other teams")))
  return(coverage)
}
data_labels <-function(df) {

  ensemble_cov <-df%>%
    filter(model=="COVIDhub-4_week_ensemble")

  data_labels <-df %>%  
    filter(model!="COVIDhub-4_week_ensemble") %>%
    mutate(test_95=ifelse(expected_cov==0.95 & observed_cov > ensemble_cov$observed_cov[ensemble_cov$expected_cov==0.95], 1, NA),
           test_80=ifelse(expected_cov==0.80 & observed_cov > ensemble_cov$observed_cov[ensemble_cov$expected_cov==0.80], 1, NA),
           test_50=ifelse(expected_cov==0.50 & observed_cov > ensemble_cov$observed_cov[ensemble_cov$expected_cov==0.50], 1, NA)) %>%
    filter(!is.na(test_95) | !is.na(test_80) | !is.na(test_50)) %>%
    group_by(model) %>%
    mutate(model_count=n()) %>%
    filter(model_count==3 & expected_cov==0.95) %>%
    mutate(model=ifelse(model=="JHU_UNC_GAS-StatMechPool", "JHU_UNC_GAS\n-StatMechPool", model)) %>%
    mutate(model=ifelse(model=="MOBS-GLEAM_COVID", "MOBS-GLEAM_\nCOVID", model)) %>%
     mutate(model=ifelse(model=="IEM_MED-CovidProject", "IEM_MED\n-CovidProject", model)) %>%
  return(data_labels)
  }
cov.plot <-function(df, df2) {
  ggplot(df) +
    geom_line(data=filter(df, bold=="F"), 
              mapping=aes(expected_cov, observed_cov, group=model, color=grouping, size=bold, alpha=0.75)) +
    geom_point(data=filter(df, bold=="F"), 
               mapping=aes(expected_cov, observed_cov, group=model, color=grouping), size=1.5) +
    geom_line(data=filter(df, bold=="T"), 
              mapping=aes(expected_cov, observed_cov, group=model, color=grouping, size=bold, alpha=0.75)) +
    geom_point(data=filter(df, bold=="T"), 
               mapping=aes(expected_cov, observed_cov, group=model, color=grouping), size=3) +
    geom_line(aes(expected_cov, expected_cov, color=grouping, group=model), linetype="dashed", size=1.5, color="black") +
    scale_size_manual(values = c("T"=1.5, "F"=0.85)) +
    scale_colour_manual(values = c("COVIDhub-baseline"="#762a83",
                                   "COVIDhub-4_week_ensemble"="#1b7837",
                                   "Other teams"="grey75"),
                        labels=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble"),
                        breaks=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble")) +
    guides(size = "none", alpha="none") +
    labs(y="Observed Coverage") +
    theme_bw()  + 
    theme(legend.position = "none", 
          axis.title.x = element_blank()) +
    geom_text_repel(data=df2,
                    mapping=aes(expected_cov, observed_cov, label=model), size=3,
                    force=1, point.padding=unit(0.5,'lines'), nudge_x =0.1,
                    segment.size= 0.05,  segment.color="grey50",
                    direction="y", #hjust = "right",
                    vjust=0.5) +
    scale_x_continuous(expand = expansion(mult = c(0.05, 0.25)))
}

# Figures
load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))

coverage_US <-cov.fig_dat(filter(dat_US_state, location=="US", sub_date <= cut_date))
labels_US <-data_labels(coverage_US)
Cov.Nat <-cov.plot(coverage_US, labels_US) +
  theme(legend.position = c(0.02, 0.88), legend.title = element_blank(),
        legend.justification = c(0, 0.5), 
        legend.margin=margin(c(0.5,1,1,1)),
        legend.key = element_rect(colour = NA, fill = NA)) +
  labs(title="A. Coverage for National Forecasts")

coverage_states <-cov.fig_dat(filter(dat_US_state, location!="US", sub_date <= cut_date))
labels_states <-data_labels(coverage_states)
Cov.State <-cov.plot(coverage_states, labels_states) +
  labs(title="B. Coverage for National State/Territory/DC Forecasts")

# Largest Counties
load(here("Data/cases_large counties_forecasts_for analysis_2022-08-18.Rdata"))
coverage_count <-cov.fig_dat(filter(dat_large_count, sub_date <= cut_date))
labels_count <-data_labels(coverage_count) 
Cov.Co <-cov.plot(coverage_count, labels_count) +
  labs(title="C. Coverage for Large County Forecasts*",
       x="Expected Coverage")

Fig3 <-annotate_figure(
  grid.arrange(Cov.Nat, Cov.State, Cov.Co, ncol=1),
  bottom = text_grob("*U.S. counties with a population of at least 93,230 people ", hjust=1, x=1, face = "italic", size = 10))

```

```{r Fig3, echo=FALSE, warning=F, message=F, fig.height=15, fig.width=6}

Fig3

```

### Figure 4. Relative WIS Across Geospatial Units

```{r Fig4, echo=FALSE, warning=F, message=F, fig.height=8, fig.width=6.5}

# All counties
relWIS_co <-function(){

  load(here("Data/WIS_all horizons_large counties.Rdata"))

  rel_WIS_large_co <-WIS_all %>%
    summarise_scores(by = c("model"),
                     metric="interval_score",
                     relative_skill = TRUE,
                     baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill) %>%
    distinct() %>%
    mutate(geo_scale="County forecasts: Size Quantile 5 (Largest)")

  load(here("Data/WIS_all horizons_non large counties_all.Rdata"))
  rel_WIS_co <-WIS_all %>%
    ungroup() %>%
    summarise_scores(by = c("model", "pop_size_quant2"),
                     metric="interval_score",
                     relative_skill = TRUE,
                     baseline = "COVIDhub-baseline") %>%
    dplyr::select(model, scaled_rel_skill, pop_size_quant2) %>%
    mutate(geo_scale=ifelse(pop_size_quant2==1, "County Forecasts: Size Quantile 1 (Smallest)",
                            ifelse(pop_size_quant2==2, "County Forecasts: Size Quantile 2",
                                   ifelse(pop_size_quant2==3, "County Forecasts: Size Quantile 3", "County Forecasts: Size Quantile 4" 
                                          #"County forecasts: Size Quantile 5 (Largest)"
                                          )))) %>%
    dplyr::select(-pop_size_quant2) %>%
    bind_rows(., rel_WIS_large_co)
}
relWIS_co <-relWIS_co()
models_include_co <-relWIS_co %>%
  filter(geo_scale=="County Forecasts: Size Quantile 1 (Smallest)") %>%
  dplyr::select(model, geo_scale) %>%
  distinct()

# Merge
models_include <-mean_WIS %>% 
  dplyr::select(-value_display, 
                -model_reorder) %>%
  spread(key,value) %>%
  drop_na()

relWIS_all <-mean_WIS %>%
  rename(scaled_rel_skill=value,
         geo_scale=key) %>%
  mutate(geo_scale=as.character(geo_scale)) %>% 
  bind_rows(., relWIS_co) %>%
  filter(model %in% models_include_co$model) %>%
  mutate(model_reorder=factor(model, levels=levels_model))

relWIS_all$geo_scale <-factor(relWIS_all$geo_scale, levels = c("County Forecasts: Size Quantile 1 (Smallest)", 
                                                               "County Forecasts: Size Quantile 2", 
                                                               "County Forecasts: Size Quantile 3", 
                                                               "County Forecasts: Size Quantile 4", 
                                                               "County forecasts: Size Quantile 5 (Largest)",
                                                               "State/Territory/DC Forecasts", "National Forecasts"))

# Colors
my_colors <-viridis(6) 
my_colors <-my_colors[-(6)] #removes the yellow color, which can be hard to see 
my_colors <-c(my_colors, "#d01c8b", "#f1b6da")

Fig4.B <-relWIS_all %>%
  mutate(model_reorder = fct_rev(model_reorder)) %>%
  filter(model!="COVIDhub-baseline") %>%
  ggplot(., aes(model_reorder, scaled_rel_skill)) +
  geom_hline(yintercept = 1, linetype="dashed") +
  geom_point(aes(color=geo_scale), size=4, alpha=0.85, 
             position = position_dodge(width = 0.2)) + 
  scale_color_manual(values=my_colors) +
  labs(y="Relative Weighted Interval Score" , 
       ) +
  theme_bw() +
      theme(axis.text.x =element_text(size=11, angle = 45, hjust=1), axis.text=axis.text_standard,
        axis.title.x = element_blank(), 
        legend.margin=margin(c(0.25,0.25,0.25,0.25)),
        legend.justification = c(0, 0.5),
        legend.key = element_rect(colour = NA, fill = NA), 
        legend.title = element_blank(),
        legend.direction = "horizontal",
        legend.position = "bottom", 
        plot.caption = plot.caption_standard) +
 scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) + guides(color=guide_legend(nrow=4, byrow=TRUE))

Fig4.B

```

### Figure 5. RelWIS per state/territory

```{r Fig5, echo=FALSE, warning=FALSE, message=FALSE, fig.height=10, fig.width=9}

# Forecasts
load(here("Data/WIS_all horizons_nat_state.Rdata"))

relWIS <-WIS_all %>%
    summarise_scores(by = c("model", "location_name"),
                     metric="interval_score",
                     relative_skill = TRUE,
                     baseline = "COVIDhub-baseline") %>%
  mutate(scaled_rel_skill=round(scaled_rel_skill, 1),
         log_WIS=log2(scaled_rel_skill),
         location_name=as.character(location_name))

Locations <-relWIS %>%
  dplyr::select(location_name) %>%
  distinct()

Models <-relWIS %>%
  dplyr::select(model) %>%
  distinct()

Locations_models <-crossing(Locations, Models) 
relWIS <- Locations_models %>%
  left_join(., relWIS) %>%
  mutate(model_reorder=factor(model, levels=levels_model))

# Reorder x and y axes
relWIS$location_name <-as.factor(relWIS$location_name)
relWIS$location_name <-relevel(relWIS$location_name, ref="National")
relWIS$location_name <-fct_rev(relWIS$location_name)

# Figure
relWIS$model_reorder <-fct_rev(relWIS$model_reorder)
ggplot(relWIS, aes(model_reorder, location_name)) +
  geom_tile(aes(fill=log_WIS)) +
  geom_text(aes(label=scaled_rel_skill), size=3) +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 0, na.value = "grey75", 
                       name = "Relative\nWeighted Interval Score", 
                       breaks = c(-1,0,1,2), 
                       labels =c("0.5", 1, 2, 3)) + 
  theme_bw() +
  labs(fill="Relative Weighted Interval Score") +
  theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.y = element_blank(),
        axis.text=axis.text_standard, plot.caption = plot.caption_standard)

```

### Figure 6. Absolute WIS and 95% coverage over time

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.show='hide'}

#options(scipen = 999)

# Function for WIS over time
WIS.ot.fig_dat <-function(df) {

  sub_date <-c(unique(df$sub_date))
  model <-c(as.character(unique(df$model)))
  model_dates <-crossing(sub_date, model)

  cast_dates <-df %>%
    dplyr::select(sub_date, date) %>%
    distinct() %>%
    left_join(model_dates, by="sub_date")

  dat <-df %>%
    full_join(cast_dates, ., by=c("model", "sub_date", "date")) %>%
    mutate(bold=ifelse(model=="COVIDhub-baseline" | model=="COVIDhub-4_week_ensemble", "T", "F")) %>%
    group_by(model, date, bold) %>%
    summarize(abWIS=mean(interval_score, na.rm=TRUE)) %>%
    ungroup() %>%
    mutate(grouping=ifelse(model=="COVIDhub-baseline", "COVIDhub-baseline", 
                           ifelse(model=="COVIDhub-4_week_ensemble", "COVIDhub-4_week_ensemble", "Other teams"))) #%>%
  return(dat)

}
cov.ot.fig_dat <-function(df) {

  sub_date <-c(unique(df$sub_date))
  model <-c(as.character(unique(df$model)))
  model_dates <-crossing(sub_date, model)

  cov <-df %>%
    group_by(model, target_end_date) %>%
    coverage(df=.) %>%
    summarize(cov.95=mean(coverage.95))

  dat <-df %>%
    dplyr::select(sub_date, target_end_date) %>%
    distinct() %>%
    left_join(., model_dates, by="sub_date") %>%
    dplyr::select(-sub_date) %>%
    left_join(., cov, by=c("model", "target_end_date")) %>%
    distinct() %>%
    ungroup() %>%
    mutate(grouping=ifelse(model=="COVIDhub-baseline", "COVIDhub-baseline", 
                           ifelse(model=="COVIDhub-4_week_ensemble", "COVIDhub-4_week_ensemble", "Other teams"))) %>%
    mutate(bold=ifelse(model=="COVIDhub-baseline" | model=="COVIDhub-4_week_ensemble", "T", "F"))

  return(dat)

}
ot.plot <-function(df, x, y) {

  ggplot() +
    geom_line(data=filter(df, bold=="F"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    geom_point(data=filter(df, bold=="F"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    geom_line(data=filter(df, bold=="T"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    geom_point(data=filter(df, bold=="T"), mapping=aes({{x}}, {{y}}, group=model, color=grouping, size=bold)) +
    scale_size_manual(values = c("T"=1.25, "F"=0.85)) +
    scale_x_date(date_labels="%b-%y", breaks = df$x, date_breaks ="3 months", date_minor_breaks ="1 month") +
    geom_vline(xintercept = as.numeric(cut_date),  linetype=4, size=1.05) +
    scale_colour_manual(name="",
                        values = c("COVIDhub-baseline"="#762a83",
                                   "COVIDhub-4_week_ensemble"="#1b7837",
                                   "Other teams"="grey80"),
                        labels=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble"),
                        breaks=c("Other teams", "COVIDhub-baseline", "COVIDhub-4_week_ensemble")) +
    guides(size = "none") + 
    theme_bw() +
    theme(legend.position = "none", 
          axis.title.x = element_blank())
}

###########################################################################
############################################################################

### WIS Over time
load(here("Data/WIS_all horizons_nat_state.Rdata"))

## National
WIS_ot_nat <-WIS.ot.fig_dat(filter(WIS_all, location_name=="National")) 
Fig6.A <-ot.plot(WIS_ot_nat, date, abWIS)  +
  theme(axis.title.x=element_blank(),
        legend.position = c(0.02, 0.80),legend.title = element_blank(),
        legend.justification = c(0, 0.5), 
        legend.margin=margin(c(0.25,0.25,0.25,0.25)),
        legend.key = element_rect(colour = NA, fill = NA)) +
  labs(title="A. Average Weighted Interval Score for National Forecasts",
       y = "Average Weighted Interval Score") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.35))) + theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5)) +
  scale_y_continuous(trans='log10')

## State/Territory/District
WIS_ot_state <-WIS.ot.fig_dat(filter(WIS_all, location_name!="National",
                                     location_name!="American Samoa"))
Fig6.B <-ot.plot(WIS_ot_state, date, abWIS) +
  labs(title = "B. Average Weighted Interval Score for State/Territory/DC Forecasts",
       y = "Average Weighted Interval Score") + theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5)) +
  scale_y_continuous(trans='log10')

## Large Counties
load(here("Data/WIS_all horizons_large counties.Rdata"))
WIS_ot_co <-WIS.ot.fig_dat(WIS_all)
Fig6.C <-ot.plot(WIS_ot_co, date, abWIS) +
  labs(title = "C. Average Weighted Interval Score for Large County Forecasts*",
       x = "Forecast target date",
       y = "Average Weighted Interval Score") + theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5)) +
  scale_y_continuous(trans='log10')

############################################################################
############################################################################

### Coverage Over time

load(here("Data/cases_nat_state_forecasts_for analysis_2022-08-18.Rdata"))
load(here("Data/cases_large counties_forecasts_for analysis_2022-08-18.Rdata"))

## National
cov_ot_nat <-cov.ot.fig_dat(filter(dat_US_state, location=="US", sub_date <= cut_date)) 
Fig6.D <-ot.plot(cov_ot_nat, target_end_date, cov.95)  +
  geom_hline(yintercept = 0.95,  linetype=6, size=1.05) +
  theme(axis.title.x=element_blank()) +  
  labs(title="D. 95% Coverage for National Forecasts",
       y = "95% Coverage") + theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5))

## State/Territory/District
cov_ot_state <-cov.ot.fig_dat(filter(dat_US_state, location!="US", sub_date <= cut_date))
Fig6.E <-ot.plot(cov_ot_state, target_end_date, cov.95) +
  geom_hline(yintercept = 0.95,  linetype=6, size=1.05) +
  labs(title = "E. 95% Coverage for State/Territory/DC Forecasts",
       y = "95% Coverage") + theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5))

## Large Counties
cov_ot_co <-cov.ot.fig_dat(filter(dat_large_count, sub_date <= cut_date))
Fig6.F <-ot.plot(cov_ot_co, target_end_date, cov.95) +
  geom_hline(yintercept = 0.95,  linetype=6, size=1.05) +
  labs(title = "F. 95% Coverage for Large County Forecasts*",
       x = "Forecast target date",
       y = "95% Coverage") +  theme(axis.text.y=element_text(angle=90, vjust=0.5, hjust=0.5))

############################################################################
############################################################################

wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
# The a label
teams_nat_st <-"Panels A, D, B and E include: LNQ-ens1, Microsoft_DeepSTIA, COVIDhub-4_week_ensemble, USC-SI_kJalpha, CU-select, LANL-GrowthRate, JHU_CSSE-DECOM, COVIDhub-trained_ensemble, COVIDhub-baseline, Karlen-pypm, BPagano-RtDriven, JHUAPL-Bucky, UVA-Ensemble, IEM_MED-CovidProject, CEID-Walk, Covid19Sim-Simulator, IowasStateLW-STEM, UCLA-SuEIR, JHU_IDD-CovidSP, RobertWalraven-ESG, MOBS-GLEAM_COVID, and CovidAnalytics-DELPHI. "

teams_co <-"Panels C and F include: LNQ-ens1, COVIDhub-4_week_ensemble, CU-select, LANL-GrowthRate, COVIDhub-trained_ensemble, COVIDhub-baseline, JHUAPL-Bucky, UVA-Ensemble, CEID-Walk, JHU_UNC_GAS-StatMechPoole, IowasStateLW-STEM, JHU_IDD-CovidSP, UMass-MechBayes, FAIR-NRAR, FRBSF_Wilson-Econometric."

############################################################################
############################################################################

Fig6 <-annotate_figure(
  grid.arrange(Fig6.A, Fig6.D,
               Fig6.B, Fig6.E,
               Fig6.C, Fig6.F,
             #Fig6.D, Fig6.H,
             ncol=2),
          bottom = text_grob("*U.S. counties with a population of at least 93,230 people \n", hjust=1, x=1, face = "italic", size = 10))

```

```{r echo=FALSE, warning=FALSE, message=FALSE}

Fig6.note1 <-annotate_figure(Fig6,
                bottom = text_grob(label = wrapper(teams_nat_st, width = 195), hjust=0, x=0.001, size = 10))

```

```{r Fig6, echo=FALSE, warning=FALSE, message=FALSE, fig.height=11, fig.width=13}

annotate_figure(Fig6.note1,
                bottom = text_grob(paste0("\n", label = wrapper(teams_co, width = 195)), hjust=0, x=0.001,
                                   size = 10))

```

### Figure 7 . Marginal mean WIS

```{r echo=FALSE, warning=F, message=F}

load(here("Data/WIS_gee.Rdata")) # data
load(here("Data/gee_for_manuscript.rda"))

team <-c(unique(as.character(WIS_gee$model))) 
dat_team <-list()
means_sds <-list()
means_sds_phase <-list()

marg_means_int <-list()
marg_means <-list()

# Model Predictions 
for(i in 1:length(team)){

  dat_team[[i]] <-as.data.frame(WIS_gee %>%
                                  filter(model==team[i]) %>%
                                  mutate(model=droplevels(model)) %>%
                                  ungroup())

  means_sds[[i]] <-dat_team[[i]] %>%
    mutate(mean_log_WIS=mean(log(interval_score)),
           sd_log_WIS=sd(log(interval_score)),
           mean_log_cases=mean(log(true_value)),
           sd_log_cases=sd(log(true_value))
    ) %>%
    dplyr::select(model, contains("mean_"), contains("sd_")) %>%
    distinct()

  # Log(cases)
  marg_means_int[[i]] <-as.data.frame(emmeans(model[[i]], ~1))
  marg_means_int[[i]] <-marg_means_int[[i]] %>%
    rename(predicted=emmean) %>%
    mutate(model=team[i],
           predicted_back=(predicted*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_lo=(asymp.LCL*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_up=(asymp.UCL*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS) %>%
    mutate(predicted_back_rescale=exp(predicted_back),
           predicted_back_lo_rescale=exp(predicted_back_lo),
           predicted_back_up_rescale=exp(predicted_back_up))

  marg_means[[i]] <-ggemmeans(model[[i]], terms=c("phase")) %>% #, "cases_log_std[-2:2]")) %>%
    rename(phase=x,
           cases=group) %>%

    mutate(model=team[i],
           predicted_back=(predicted*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_lo=(conf.low*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_up=(conf.high*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS) %>%

    mutate(predicted_back_rescale=exp(predicted_back),
           predicted_back_lo_rescale=exp(predicted_back_lo),
           predicted_back_up_rescale=exp(predicted_back_up)) 
}
marg_means_int <-do.call(rbind.data.frame, marg_means_int)
marg_means <-do.call(rbind.data.frame, marg_means)

marg_means_int <-marg_means_int %>% 
  dplyr::select(-`1`) %>%
  mutate(model=as.factor(model),
         model=fct_reorder(model, predicted_back),
         phase="Mean across phases") 
levels <-levels(marg_means_int$model)

marg_means <-marg_means %>%

  ## Order by increasing error, on average
  mutate(model=as.factor(model),
         model=factor(model, levels = levels)) %>%
  mutate(phase=as.factor(ifelse(phase=="increasing", "Increasing",
                                    ifelse(phase=="decreasing", "Decreasing",
                                           ifelse(phase=="peak", "Peak",
                                                  ifelse(phase=="nadir", "Nadir", "Mean across phases"))))),
         phase=factor(phase, 
                          levels = c("Mean across phases", "Increasing", "Peak", "Decreasing",  "Nadir"))
  ) %>%
  ungroup()

baseline <-filter(marg_means_int, model=="COVIDhub-baseline")

# Figure
Fig7A <-marg_means_int %>%
  ggplot(.) +
  geom_point(aes(model, predicted_back_rescale)) +
  geom_linerange(aes(model, ymin=predicted_back_lo_rescale, ymax=predicted_back_up_rescale)) +
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_lo_rescale), color="red", linetype="dashed") + 
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_up_rescale), color="red", linetype="dashed") +
  labs( y="Marginal mean\n Weighted Interval Score", 
        title="A.") +
  theme_bw() +
  theme(axis.title.y=axis.title.y_standard, axis.text.y=axis.text_standard,
        legend.position = "bottom", legend.title = element_blank(), legend.text=legend.text_standard,
        strip.text=strip.text_standard, strip.background=strip.background_standard,
        axis.title.x = element_blank(),
        #axis.text.x = element_text(angle = 45, hjust=1),
        axis.text.x = element_blank())

```

```{r echo=FALSE, warning=F, message=F}

Fig7B <-marg_means %>%
  mutate(scaled_pred=predicted_back/baseline$predicted_back,
         SE_scaled=(predicted_back_up-predicted_back_lo)/3.92,
         prob_base=round(pnorm(baseline$predicted_back,
                               predicted_back, SE_scaled), 5),
         prob_label=ifelse(prob_base >=0.9, "*", NA),
         phase=fct_rev(phase)) %>%
  ggplot(.) +
  geom_tile(aes(model, phase, fill=scaled_pred)) +
  geom_text(aes(model, phase, label=prob_label), size=4) +
  scale_fill_gradient2(low = "#7fcdbb", high = "#fecc5c", midpoint = 1.0,
                       name="Marginal mean Weighted Interval Score  \nscaled to the marginal mean  \nof COVIDhub-baseline model  \nacross all phases",
                       guide = guide_colourbar(barheight = 1.5)) +
  labs(y="", title="B.",
       caption="*80% CI is lower than baseline") + 
  theme_bw() +
  theme(legend.position = "bottom", legend.title.align=0.5, legend.box.just = "center", 
    plot.caption = plot.caption_standard,
    strip.text=strip.text_standard, strip.background=strip.background_standard,
    panel.border = element_rect(colour = "black", fill = NA),
    axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1), axis.text=axis.text_standard,
    axis.ticks.y=element_blank())

```

```{r Fig7, echo=FALSE, warning=F, message=F, fig.height=8.5, fig.width=8.5}

plot_grid(Fig7A, Fig7B, rel_heights  = c(0.5, 1), axis = "l",
          align="v", nrow = 2)

```