### 2023-04-18 Code/Supplement 2_Revisions.Rmd

---
title: "Supplement 2: Team Specific GEEs  & Exploration of Revision Impact"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(here)
library(tidyverse)
library(covidcast)
library(epiprocess)

library(geepack)
library(splines)

library(ggeffects)
library(emmeans)

library(gridExtra)
library(grid)
library(cowplot)

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size = 12, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=10, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

## Read in data used in analysis
main_dat <-covidcast_signal(data_source = "jhu-csse",
                      signal = "confirmed_incidence_num",
                      start_day = "2020-07-26",
                      end_day = "2022-01-15",
                      geo_type = "state",
                      as_of = "2022-04-02") %>%
  dplyr::select(geo_value, time_value, cases_now=value) %>%   
  filter(geo_value!="as", geo_value!="gu", geo_value!="mp",
         geo_value!="pr", geo_value!="vi"#, geo_value!='dc'
         ) %>% 
  group_by(geo_value)

dates <-seq.Date(as.Date("2020-08-01"), as.Date("2022-01-15"), by="week")
hx_dat <-list()
for(i in 1:length(dates)){
  hx_dat[[i]] <-covidcast_signal(data_source = "jhu-csse",
                      signal = "confirmed_incidence_num",
                      start_day =dates[[i]] - 6,
                      end_day = dates[[i]], 
                      geo_type = "state",
                      as_of = dates[[i]] +1) %>%
  dplyr::select(geo_value, time_value, cases_last=value) %>%   
  filter(geo_value!="as", geo_value!="gu", geo_value!="mp",
         geo_value!="pr", geo_value!="vi"#, geo_value!='dc'
         ) %>% 
  mutate(wk = as.Date(time_value, "%m/%d/%y"), 
         wk = as.Date(cut(wk,"week", start.on.monday = FALSE)) + 6)  
}
hx_dats_all <-do.call(rbind, hx_dat)

cases_all <-main_dat %>%
  left_join(., hx_dats_all, by=c("geo_value", "time_value")) %>%
  mutate(revision=ifelse(cases_now==cases_last, 0, 1)) %>%
  filter(revision==1) %>%
  dplyr::select(geo_value, wk, revision) %>%
  distinct()

## rename locations based on state names
states <-read.csv(here("Data/locations.csv")) %>%
  rename(fips = location) %>%
  mutate(fips = as.character(fips)) %>%
  filter(nchar(fips)==2) %>%
  dplyr::select(abbreviation, location_name)

cases_all <-cases_all %>%
  rename(abbreviation=geo_value,
         date=wk) %>%
  mutate(abbreviation=toupper(abbreviation)) %>%
  left_join(., states, by="abbreviation") %>%
  distinct() %>%
  ungroup() %>%
  dplyr::select(-abbreviation) %>%
  mutate(location_name=as.factor(location_name))

## Read in data
load(here("Data/WIS_gee.Rdata"))

WIS_rev <-WIS_gee %>%
   arrange(location_name, date) %>%
   left_join(., cases_all, by=c("location_name", "date")) %>%
   mutate(revision=ifelse(is.na(revision), 0, revision)) %>%
   group_by(model, location_name, target) %>%
   ungroup() %>%
   arrange(location_name, date) %>%
   filter(revision==0) %>%
   mutate(location_name=droplevels(location_name))

```

```{r echo=FALSE, warning=FALSE, message=FALSE}

team <-c(as.character(unique(WIS_rev$model)))
dat_team_rev<-list()

model_rev <-list()

for(i in 1:length(team)){

  dat_team_rev[[i]] <-WIS_rev %>%
    filter(model==team[i]) %>%
    mutate(model=droplevels(model))

  model_rev[[i]] <-geeglm(formula =  WIS_log_std ~ ns(cases_log_std, df=2) + phase + target,
                      family = gaussian, data = dat_team_rev[[i]], id = location_name, wave=date,
                      corstr = "independence")  
}

```

```{r echo=FALSE, warning=FALSE, message=FALSE}

team <-c(unique(as.character(WIS_rev$model))) 
dat_team_rev <-list()
means_sds <-list()

marg_means_int <-list()
marg_means <-list()

# Model Predictions 
for(i in 1:length(team)){

  dat_team_rev[[i]] <-as.data.frame(WIS_rev %>%
                                  filter(model==team[i]) %>%
                                  mutate(model=droplevels(model)) %>%
                                  ungroup())

  means_sds[[i]] <-dat_team_rev[[i]] %>%
    mutate(mean_log_WIS=mean(log(interval_score)),
           sd_log_WIS=sd(log(interval_score)),
           mean_log_cases=mean(log(true_value)),
           sd_log_cases=sd(log(true_value))
    ) %>%
    dplyr::select(model, contains("mean_"), contains("sd_")) %>%
    distinct()

  # Intercept
  marg_means_int[[i]] <-as.data.frame(emmeans(model_rev[[i]], ~1))
  marg_means_int[[i]] <-marg_means_int[[i]] %>%
    rename(predicted=emmean) %>%
    mutate(model=team[i],
           predicted_back=(predicted*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_lo=(asymp.LCL*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_up=(asymp.UCL*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS) %>%
    mutate(predicted_back_rescale=exp(predicted_back),
           predicted_back_lo_rescale=exp(predicted_back_lo),
           predicted_back_up_rescale=exp(predicted_back_up))

  #Phases
  marg_means[[i]] <-ggemmeans(model_rev[[i]], terms=c("phase")) %>% #, "cases_log_std[-2:2]")) %>%
    rename(phase=x,
           cases=group) %>%

    mutate(model=team[i],
           predicted_back=(predicted*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_lo=(conf.low*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS,
           predicted_back_up=(conf.high*means_sds[[i]]$sd_log_WIS) + means_sds[[i]]$mean_log_WIS) %>%
    mutate(predicted_back_rescale=exp(predicted_back),
           predicted_back_lo_rescale=exp(predicted_back_lo),
           predicted_back_up_rescale=exp(predicted_back_up)) 
}
marg_means_int <-do.call(rbind.data.frame, marg_means_int)
marg_means <-do.call(rbind.data.frame, marg_means)

marg_means_int <-marg_means_int %>% 
  dplyr::select(-`1`) %>%
  mutate(model=as.factor(model),
         model=fct_reorder(model, predicted_back),
         phase="Mean across phases") 
levels <-levels(marg_means_int$model)

marg_means <-marg_means %>%

  ## Order by increasing error, on average
  mutate(model=as.factor(model),
         model=factor(model, levels = levels)) %>%
  mutate(phase=as.factor(ifelse(phase=="increasing", "Increasing",
                                    ifelse(phase=="decreasing", "Decreasing",
                                           ifelse(phase=="peak", "Peak",
                                                  ifelse(phase=="nadir", "Nadir", "Mean across phases"))))),
         phase=factor(phase, 
                          levels = c("Mean across phases", "Increasing", "Peak", "Decreasing",  "Nadir"))
  ) %>%
  ungroup()

baseline <-filter(marg_means_int, model=="COVIDhub-baseline")

# Figure
FigS2.2 <-marg_means_int %>%
  ggplot(.) +
  geom_point(aes(model, predicted_back_rescale)) +
  geom_linerange(aes(model, ymin=predicted_back_lo_rescale, ymax=predicted_back_up_rescale)) +
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_lo_rescale), color="red", linetype="dashed") + 
  geom_hline(data=baseline, mapping=aes(yintercept=baseline$predicted_back_up_rescale), color="red", linetype="dashed") +
  labs( y="Marginal Mean\n Weighted Interval Score", 
        title="A.") +
  theme_bw() +
  theme(axis.title.y=axis.title.y_standard, axis.text=axis.text_standard,
        legend.position = "bottom", legend.title = element_blank(), legend.text=legend.text_standard,
        strip.text=strip.text_standard, strip.background=strip.background_standard,
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))

Fig6B <-marg_means %>%
  mutate(scaled_pred=predicted_back/baseline$predicted_back,
         SE_scaled=(predicted_back_up-predicted_back_lo)/3.92,
         prob_base=round(pnorm(baseline$predicted_back,
                               predicted_back, SE_scaled), 5),
         prob_label=ifelse(prob_base >=0.9, "*", NA),
         phase=fct_rev(phase)) %>%
  ggplot(.) +
  geom_tile(aes(model, phase, fill=scaled_pred)) +
  geom_text(aes(model, phase, label=prob_label), size=4) +
  scale_fill_gradient2(low = "#7fcdbb", high = "#fecc5c", midpoint = 1.0,
                       name="Marginal mean Weighted Interval Score  \nscaled to the marginal mean  \nof COVIDhub-baseline model  \nacross all phases",
                       guide = guide_colourbar(barheight = 1.5)) +
  labs(y="", title="B.",
       caption="*80% CI is lower than baseline") + 
  theme_bw() +
  theme(#axis.title.y=axis.title.y_standard,
    legend.position = "bottom", legend.title.align=0.5, legend.box.just = "center", 
    plot.caption = plot.caption_standard,
    strip.text=strip.text_standard, strip.background=strip.background_standard,
    panel.border = element_rect(colour = "black", fill = NA),
    axis.title.x = element_blank(), axis.text.x = element_text(angle = 45, hjust=1), axis.text=axis.text_standard,
    axis.ticks.y=element_blank())#, axis.text.y=element_blank())

```

```{r Fig S2.2, echo=FALSE, warning=F, message=F, fig.height=8.5, fig.width=8.5}

FigS2.2_updated <-FigS2.2 + theme(axis.text.x = element_blank())

plot_grid(FigS2.2_updated, Fig6B, rel_heights  = c(0.5, 1), axis = "l",
          align="v", nrow = 2)

```