### 2024-01-10 Code/Supplement 4_Spatial Correlations.Rmd
add caption to figure
per journal review request re copyright

---
title: "Supplement 4: Spatial Correlations"
author: "VKL"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning=FALSE,message=FALSE)

library(here)
library(tidyverse)

library(scoringutils)
library(ggthemes)

library(sf)
library(spdep)

## Standardized formatting
strip.background_standard <-element_blank()
strip.text_standard <-element_text(size = 12, face='bold')
legend.text_standard <-element_text(size=12)
plot.caption_standard <-element_text(size=10, face = "italic")
axis.title.y_standard <-element_text(margin = margin(1, 10, 1, 1), size=12)
axis.text_standard <-element_text(size=11)

## Read in data used in analysis
load("./Data/WIS_all horizons_nat_state.rdata")
WIS_st <-WIS_all %>%
  ungroup() %>%
  filter(location_name!="National") %>%
  summarise_scores(by = c("model", "location_name"),
                   metric="interval_score",
                   relative_skill = TRUE,
                   baseline = "COVIDhub-baseline") %>%
  mutate(location_name=tolower(location_name)) %>%
  filter(model!="COVIDhub-baseline")

```

```{r echo=FALSE, warning=FALSE, message=FALSE}

## Run Moran's I for spatial correlation of WIS estimates (states only)

set.seed(2354)

WIS_st$location_name <-str_to_title(WIS_st$location_name)
exclude <-c("Alaska", "Hawaii", "Puerto Rico")
coords <-tigris::states(class="sf", cb=TRUE) %>%
  mutate(NAME=str_to_title(NAME)) %>%
  dplyr::select(geometry, NAME) %>%
  filter(!NAME %in% exclude)

mods <-sort(unique(WIS_st$model))
MI <-list()

for(i in 1:length(mods)){
  WIS_st_sub <-WIS_st %>%
    dplyr::select(scaled_rel_skill, model, location_name) %>%
    filter(model==mods[i])

  s <-coords %>%
    left_join(., WIS_st_sub, by=c("NAME"="location_name")) %>%
    drop_na() %>%
    dplyr::select(-model)

  nb <-poly2nb(s, queen=TRUE) # Define the neighbors (use queen case)
  lw <-nb %>% nb2listw(., style="W", zero.policy=TRUE) # Compute the neighboring average WIS

  MI[[i]] <- moran.mc(s$scaled_rel_skill, lw, nsim=1000, alternative="greater") # Run the MC simulation version of the Moran's I test

}

MI_stat <-list()
MI_p <-list()
for(i in 1:length(mods)){
  MI_stat[[i]] <-MI[[i]]$statistic
  MI_p[[i]] <-MI[[i]]$p.value
}
stats <-round(do.call(c, MI_stat), 2)
ps <-round(do.call(c, MI_p), 3)

MI_results <-bind_cols(mods, stats, ps) %>%
  rename(model=1,
         morans=2, 
         p_value=3)

```

```{r Fig S4.1, echo=FALSE, warning=F, message=F, fig.height=8.5, fig.width=11}

### National WIS (for reordering teams in plot)
WIS_nat <-WIS_all %>%
  filter(location_name=="National") %>% 
  summarise_scores(by = c("model"),
                   metric="interval_score",
                   relative_skill = TRUE,
                   baseline = "COVIDhub-baseline") %>%
  left_join(., MI_results, by=c("model")) %>%
  filter(model!="COVIDhub-baseline") %>%
  mutate(order=paste0(model, "\nMoran's I: ", morans, " (p-value ", p_value, ")"),
         order=as.factor(ifelse(scaled_rel_skill < 1.0, paste0(model, "*\nMoran's I: ", morans, " (p-value ", p_value, ")"), order)),
         order=fct_reorder(order, scaled_rel_skill))
WIS_levels <-levels(WIS_nat$order) 
mods_better <-WIS_nat %>%
  filter(scaled_rel_skill < 1.0)

### Update map with Moran I's and reorder
us <-map_data("state")
WIS_map <-coords %>%
  left_join(., WIS_st, by=c("NAME"="location_name")) %>%
  left_join(., MI_results, by=c("model")) %>%
  mutate(model_group=ifelse(model %in% mods_better$model, "National rWIS < 1.0", "National rWIS > 1.0"), 
         model=ifelse(model %in% mods_better$model, paste0(model, "*"), model), 
         order=paste0(model, "\nMoran's I: ", morans, " (p-value ", p_value, ")"),
         order_fact=factor(order, levels=WIS_levels)) %>% drop_na()

WIS_map %>%
  ggplot(., aes(fill = log(scaled_rel_skill))) +
  geom_sf(color = "grey65") + 
  facet_wrap(~order_fact) +
  theme_void() +
  scale_fill_gradient2(low = "steelblue", high = "red", midpoint = 0, na.value = "grey75", 
                       name = "State-Level Relative\nWeighted Interval Score", 
                       breaks = c(-1,0,1,2), 
                       labels =c("0.5", 1, 2, 3)) + 
  labs(caption="*National rWIS < 1.0\nThe base layer for this map was created with ggplot2 (https://ggplot2.tidyverse.org/reference/map_data.html).") +
  theme(plot.margin = unit(c(0,0,0,0), "cm"),
        strip.background=element_blank(), strip.text=element_text(size = 10, face='bold'),
        legend.justification = c(1,0), legend.position=c(0.90, 0.10),
        legend.direction = "horizontal")

```